---
subtitle: "TMA4268 Statistical Learning V2022"
title: "Compulsory exercise 2: Group 18"
author: "Thomas Rødland, Erlend Lokna"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  # html_document
  pdf_document
---

```{r setup, eval=TRUE,echo=FALSE, message=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3,fig.align = "center")
```


```{r, eval=TRUE,echo=FALSE, message=FALSE}
library("ggplot2")
library("tidyverse")
library("GGally")
library("MASS")
library("caret")
library("glmnet")
library("pls")
library("gam")
library("e1071")
library("tree")
library("randomForest")
library("ggfortify")
library("leaps")
```

```{r, eval=TRUE}
set.seed(1)
boston <- scale(Boston, center=T, scale=T)

# split into training and rest sets
train.ind = sample(1:nrow(boston), 0.8 * nrow(boston))
boston.train = data.frame(boston[train.ind, ])
boston.test = data.frame(boston[-train.ind, ])
```

# Problem 1

## a)
```{r, eval=TRUE}
#forward subset selection:
regfit_fwd.full = regsubsets(medv~., data = boston.train, nvmax=13, method = "forward")
regfit_fwd.four = regsubsets(medv~., data = boston.train, nvmax=4, method = "forward")

#plotting the R-squared vs number of predictors
plot(summary(regfit_fwd.four)$rsq, xlab='No. of Variables',ylab='R^2',type='b')
```

## b)
```{r, eval=TRUE}
#the best four predictors:
coef(regfit_fwd.four, 4)
```
## c)
```{r, eval=TRUE}
#K-fold cross-validation (k=5) 
#setup:
df.X.std <- scale(dplyr::select(Boston, -medv))       
X.train <- as.matrix(df.X.std)[train.ind,] #converting to matrix. 
X.test <- as.matrix(df.X.std)[-train.ind,]
Y.train <- Boston[train.ind, "medv"]
Y.train = as.matrix(Y.train)
Y.test <- Boston[-train.ind, "medv"]
Y.test = as.matrix(Y.test)

#calculation using cv.glmnet():
kf5_cv = cv.glmnet(x=X.train, y=Y.train, alpha = 1, nfolds = 5)
```

```{r, eval=TRUE}
#plot:
plot(kf5_cv)
```
Lambda corresponding to minimal Mean-Squared error:
```{r, eval=TRUE}
kf5_cv$lambda.min
```
coefficients:
```{r, eval=TRUE}
coef(kf5_cv, s=kf5_cv$lambda.min)
```
## d)
1. True
2. False
3.
4. True

# Problem 2

```{r, eval=TRUE}
set.seed(1)

# load a synthetic dataset
id <- "1CWZYfrLOrFdrIZ6Hv73e3xxt0SFgU4Ph"  # google file ID
synthetic <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))

# split into training and test sets
train.ind = sample(1:nrow(synthetic), 0.8 * nrow(synthetic))
synthetic.train = data.frame(synthetic[train.ind, ])
synthetic.test = data.frame(synthetic[-train.ind, ])

# show head(..)  Y: response variable; X: predictor variable
head(synthetic)
```

### a)

Fitting PCR and PSLR on the synthetic.train data set:

```{r, eval=TRUE}
pcr_model <- pcr(Y~., data = synthetic.train, scale = TRUE, validation = "CV")
pslr_model <- plsr(Y~., data=synthetic.train, scale = TRUE, validation = "CV")
```

```{r, eval=TRUE}
validationplot(pcr_model, val.type="MSEP", main="PCR")
```

```{r, eval=TRUE}
validationplot(pslr_model, val.type="MSEP", main="PSLR")
```

### b)

We can clearly see that the PSLR method shrinks the MSEP quicker for fewer components when comparing to the PCR method. PCR is a unsupervised method while PSLR is supervised. 


# Problem 3

## a)
1.
2.
3. False (the extra term makes it less wiggly not more smooth?)
4. True (increase in k -> higher bias, lower variance ?)

## b)
using gam function to create a adaptive model:
```{r, eval=TRUE}
adpt_mod <- gam(medv ~rm  + s(ptratio, df=3) + s(lstat, df=2), data=boston.train)
```

plotting results:
```{r, eval=TRUE}
par(mfrow=c(1,3)) #to partition the Plotting Window
plot(adpt_mod, se=TRUE)
```

